---
title: "PSTAT 277B: Synthetic Data Application"
author: "Isaiah Katz"
date: "Winter 2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
library(tidyverse)
library(MASS)
library(urca)
library(tsDyn)
library(vars)
library(tseries)

library(viridis)
library(grid)
library(gridExtra)
```

### Example 1: white noise process 

Testing Johansen procedure on multivariate series composed of I(1), but non-cointegrated components. Components simulated as AR(1) white noise with $\phi_1 = 1$. 

```{r, EXAMPLE 1 I(1) NO COINTEGRATION, fig.align="center", out.width="50%"}
## Example 1: multivariate time series composed of independent I(1) processes
set.seed(1729)

## parameter specification 
d <- 3; n <- 500
epsilon.t <- mvrnorm(n, rep(0, d), Sigma = diag(1,d))

## toy model 
yt.ex1 <- apply(epsilon.t, 2, cumsum) 
colnames(yt.ex1) <- c("y1", "y2", "y3")

ts.plot(yt.ex1, lty = c(1:d), main="3-Dimensional I(1) Series Without Cointegration", 
        col = rainbow(d))
legend(0, -10, c("y1","y2", "y3"), lty=c(1:d), col = rainbow(d))
```

```{r, EXAMPLE 1 COINTEGRATION CHECK}
## test for presence of unit roots; all processes nonstationary 
adf.test(yt.ex1[,1], k = 1)
adf.test(yt.ex1[,2], k = 1)
adf.test(yt.ex1[,3], k = 1)

## confirm all processes are I(1) 
adf.test(diff(yt.ex1[,1]))
adf.test(diff(yt.ex1[,2]))
adf.test(diff(yt.ex1[,3]))

## test cointegration rank with Johansen methodology 
p.ex1 <- VARselect(yt.ex1)
p.ex1$selection

## trace and eigenvalue statistic using urca -> ca.jo() 
ex1.eigen.stat <- ca.jo(yt.ex1, type = "eigen", ecdet = "none", K = 2, 
                        spec = "transitory")
summary(ex1.eigen.stat)

ex1.trace.stat <- ca.jo(yt.ex1, type = "trace", ecdet = "none", K = 2, 
                        spec = "transitory")
summary(ex1.trace.stat)

## trace and eigenvalue statistic using tsDyn -> VECM() and rank.test()
ex1.vecm <- VECM(yt.ex1, lag = 1, estim = "ML", include = "const")
ex1.combo.stat <- rank.test(ex1.vecm)
summary(ex1.combo.stat)
summary(ex1.vecm)
```

Johansen test suggests no cointegration present, as expected.  

### Example 2: Bivariate cointegrated VAR(p) process 

Testing Johansen procedure on simulated bivariate VAR(p) process.

```{r, EXAMPLE 2 BIVARIATE COINTEGRATED, fig.align="center", out.width="50%"}
## Example 2: bivariate cointegrated series 
set.seed(1729) 

## parameter specification
d <- 2; n <- 500
epsilon.t <- mvrnorm(n, mu = rep(0, 3), Sigma = diag(1, 3))
c.trend <- cumsum(epsilon.t[,3])
e1 <- arima.sim(list(ar = 0.5), n, innov = epsilon.t[,1])
e2 <- arima.sim(list(ar = 0.5), n, innov = epsilon.t[,2])

## cointegrating vector
beta.ex2 <- c(1, -2) 

## model construction 
yt.ex2.1 <- -beta.ex2[2] * c.trend + e1
yt.ex2.2 <- c.trend + e2
yt.ex2 <- cbind(yt.ex2.1, yt.ex2.2)
  
linecolors <- viridis(20)[c(5,15)]
ts.plot(yt.ex2, lty = c(1:d), main = "Bivariate VAR(p) Time Series", 
        col = linecolors)
legend(0, -10, c("y1","y2"), lty=c(1:d), col = linecolors)
```

```{r, EXAMPLE 2 COINTEGRATION CHECK}
## test for presence of unit roots; all processes nonstationary 
adf.test(yt.ex2[,1], k = 1)
adf.test(yt.ex2[,2], k = 1)

## confirm all processes are I(1)
adf.test(diff(yt.ex2[,1]), k = 1)
adf.test(diff(yt.ex2[,2]), k = 1)

# test cointegration rank with Johansen methodology 
p.ex2 <- VARselect(yt.ex2)
p.ex2$selection

## trace and eigenvalue statistic using urca -> ca.jo() 
ex2.eigen.stat <- ca.jo(yt.ex2, type = "eigen", ecdet = "none", K = 2, 
                        spec = "transitory")
summary(ex2.eigen.stat)

ex2.trace.stat <- ca.jo(yt.ex2, type = "trace", ecdet = "none", K = 2, 
                        spec = "transitory")
summary(ex2.trace.stat)

## trace and eigenvalue statistic using tsDyn -> VECM() and rank.test()
ex2.vecm <- VECM(yt.ex2, lag = 1, estim = "ML", include = "const")
ex2.combo.stat <- rank.test(ex2.vecm)
summary(ex2.combo.stat)
summary(ex2.vecm)

## verify estimated cointegration vector produces I(0) series   
coint.vec.ex2 <- ex2.trace.stat@V[,1]
zt.ex2 <- yt.ex2[,1]*coint.vec.ex2[1] + yt.ex2[,2]*coint.vec.ex2[2]
ur.df(zt.ex2, selectlags = "AIC")  # suggests stationarity 
```

Correctly estimated cointegration order 1; also nearly estimated cointegrating vector. Estimated cointegrating relationship yields stationary series. 

### Example 3: Trivariate Cointegrated VAR(p) Process (1 relation)

Testing Johansen procedure on simulated trivariate VAR(p) process. Examples (3) and (4) test cointegration in series with cointegration rank 1 (example 3) and cointegration rank 2 (example 4). 

```{r, EXAMPLE 3 TRIVARIATE COINTEGRATED RANK 1, fig.align="center", out.width="50%"}
## Example 3: trivariate cointegrated series, cointegration rank 1 
set.seed(1729)

## parameter specification 
n <- 500; d <- 3
epsilon.t <- mvrnorm(n, mu = rep(0, d), Sigma = diag(0.75, d))
e1 <- arima.sim(list(ar = 0.5), n, innov = epsilon.t[,1])
e2 <- arima.sim(list(ar = 0.5, -0.5), n, innov = epsilon.t[,2])
e3 <- arima.sim(list(ar = 0.5, -0.5), n, innov = epsilon.t[,3])

## cointegrating vector   
beta.ex3 <- c(1, -2, -3)

# model construction
yt.ex3.2 <- cumsum(e2)
yt.ex3.3 <- cumsum(e3)
yt.ex3.1 <- -beta.ex3[2]*yt.ex3.2 -beta.ex3[3]*yt.ex3.3 + e1
yt.ex3 <- cbind(yt.ex3.1, yt.ex3.2, yt.ex3.3)

ts.plot(yt.ex3, lty = c(1:d), 
        main = "Trivariate VAR(p): Cointegrated Rank 1", 
        col = rainbow(d))
legend(0, -50, c("y1","y2", "y3"), lty=c(1:3), col = rainbow(d))
```

```{r, EXAMPLE 3 COINTEGRATION CHECK}
## test for presence of unit roots; all processes nonstationary 
adf.test(yt.ex3[,1], k = 1)
adf.test(yt.ex3[,2], k = 1)
adf.test(yt.ex3[,3], k = 1)

## confirm all processes are I(1)
adf.test(diff(yt.ex3[,1]), k = 1)
adf.test(diff(yt.ex3[,2]), k = 1)
adf.test(diff(yt.ex3[,3]), k = 1)

# test cointegration rank with Johansen methodology 
p.ex3 <- VARselect(yt.ex3)
p.ex3$selection

## trace and eigenvalue statistic using urca -> ca.jo() 
ex3.eigen.stat <- ca.jo(yt.ex3, type = "eigen", ecdet = "none", K = 2, 
                        spec = "transitory")
summary(ex3.eigen.stat)

ex3.trace.stat <- ca.jo(yt.ex3, type = "trace", ecdet = "none", K = 2, 
                        spec = "transitory")
summary(ex3.trace.stat)

## trace and eigenvalue statistic using tsDyn -> VECM() and rank.test()
ex3.vecm <- VECM(yt.ex3, lag = 1, estim = "ML", include = "const")
ex3.combo.stat <- rank.test(ex3.vecm)
summary(ex3.combo.stat)
summary(ex3.vecm)

## verify estimated cointegration vector produces I(0) series 
coint.vec.ex3 <- ex3.trace.stat@V[,1]
zt.ex3 <- yt.ex3[,1]*coint.vec.ex3[1] + 
  yt.ex3[,2]*coint.vec.ex3[2] + yt.ex3[,3]*coint.vec.ex3[3]
ur.df(zt.ex3, selectlags = "AIC")
```

Correctly estimates cointegration order 1. Estimated cointegrating vector yields stationary series as shown by ADF statistic. 

### Example 4: Trivariate Cointegrated VAR(p) Process (2 relations)

```{r, EXAMPLE 4 TRIVARIATE COINTEGRATED RANK 2, fig.align="center", out.width="50%"}
## Example 4: trivariate cointegrated series, cointegration rank 2
set.seed(1729)

## parameter specification 
n <- 500; d <- 3
epsilon.t <- mvrnorm(n, mu = rep(0, d), Sigma = diag(1, d))
e1 <- arima.sim(list(ar = -0.5), n, innov = epsilon.t[,1])
e2 <- arima.sim(list(ar = -0.25), n, innov = epsilon.t[,2])
e3 <- arima.sim(list(ar = -0.75), n, innov = epsilon.t[,3])

## cointegrating vector   
beta.ex4.1 <- c(3, 0, 1)
beta.ex4.2 <- c(0, -2, 1)

## model construction
yt.ex4.3 <- cumsum(e3)
yt.ex4.1 <- -beta.ex4.1[1]*yt.ex4.3 + e1
yt.ex4.2 <- -beta.ex4.2[2]*yt.ex4.3 + e2
yt.ex4 <- cbind(yt.ex4.1, yt.ex4.2, yt.ex4.3)

ts.plot(yt.ex4, lty = c(1:d), 
        main = "Trivariate VAR(p): Cointegrated Rank 2", 
        col = rainbow(d))
legend(0, -10, c("y1","y2", "y3"), lty=c(1:3), col = rainbow(d))
```

```{r, EXAMPLE 4 COINTEGRATION CHECK}
## test for presence of unit roots; all processes nonstationary 
adf.test(yt.ex4[,1], k = 1)
adf.test(yt.ex4[,2], k = 1)
adf.test(yt.ex4[,3], k = 1)

## confirm all processes are I(1)
adf.test(diff(yt.ex4[,1]), k = 1)
adf.test(diff(yt.ex4[,2]), k = 1)
adf.test(diff(yt.ex4[,3]), k = 1)

# test cointegration rank with Johansen methodology 
p.ex4 <- VARselect(yt.ex4)
p.ex4$selection

## trace and eigenvalue statistic using urca -> ca.jo() 
ex4.eigen.stat <- ca.jo(yt.ex4, type = "eigen", ecdet = "none", K = 2, 
                        spec = "transitory")
summary(ex4.eigen.stat)

ex4.trace.stat <- ca.jo(yt.ex4, type = "trace", ecdet = "none", K = 2, 
                        spec = "transitory")
summary(ex4.trace.stat)

## trace and eigenvalue statistic using tsDyn -> VECM() and rank.test()
ex4.vecm <- VECM(yt.ex4, lag = 1, estim = "ML", include = "const")
ex4.combo.stat <- rank.test(ex4.vecm)
summary(ex4.combo.stat)

ex4.vecm.correct <- VECM(yt.ex4, r = 2, lag = 1, estim = "ML", include = "const")
ex4.coint.vec <- ex4.vecm.correct$model.specific$coint
summary(ex4.vecm.correct)

## verify estimated cointegration vector produces I(0) series 
coint.vec.1.ex4 <- ex4.coint.vec[,1]
coint.vec.2.ex4 <- ex4.coint.vec[,2]
zt.ex4.1 <- yt.ex4[,1]*coint.vec.1.ex4[1] + yt.ex4[,2]*coint.vec.1.ex4[2] + 
  yt.ex4[,3]*coint.vec.1.ex4[3]
zt.ex4.2 <- yt.ex4[,1]*coint.vec.2.ex4[1] + yt.ex4[,2]*coint.vec.2.ex4[2] + 
  yt.ex4[,3]*coint.vec.2.ex4[3]

ur.df(zt.ex4.1, selectlags = "AIC")
ur.df(zt.ex4.2, selectlags = "AIC")
```

Correctly estimates cointegration order 2; also reasonably accurately estimates cointegration vector. Note that VECM specification from `tsDyn` will not automatically select an appropriate cointegration rank; must be tested (i.e., via `rank.test`) and then re-fit. 

### Example 5: Trivariate Cointegrated VAR(p) With Regime Switching

Testing Johansen procedure on series with regime shift in cointegration rank and cointegration relations using a rolling window.

```{r, EXAMPLE 5: COINTEGRATION WITH REGIME SHIFTS}
## Example 5: trivariate cointegrated series, varying cointegration relations
set.seed(1729)

## Parameter specification
n <- 600; d <- 3
epsilon.t <- mvrnorm(n, mu = rep(0, d), Sigma = diag(0.75, d))
e1 <- arima.sim(list(ar = 0.5), n, innov = epsilon.t[,1])
e2 <- arima.sim(list(ar = 0.5, -0.5), n, innov = epsilon.t[,2])
e3 <- arima.sim(list(ar = 0.5, -0.5), n, innov = epsilon.t[,3])

## cointegrating vectors 
beta.ex5.r1 <- c(1, 2, 3)
beta.ex5.r2.1 <- c(0.5, 0, 1)
beta.ex5.r2.2 <- c(0, -0.5, 1)
beta.ex5.r3 <- c(1, 2, 3)

## model construction: regime 1
yt.ex5.2.r1 <- cumsum(e2[1:200])
yt.ex5.3.r1 <- cumsum(e3[1:200])
yt.ex5.1.r1 <- -beta.ex5.r1[2]*yt.ex5.2.r1 -beta.ex5.r1[3]*yt.ex5.3.r1 + e1[1:200]
yt.ex5.r1 <- cbind(yt.ex5.1.r1, yt.ex5.2.r1, yt.ex5.3.r1)

## model construction: regime 2
yt.ex5.3.r2 <- sum(e3[1:200]) + cumsum(e3[201:400])
yt.ex5.1.r2 <- -beta.ex5.r2.1[1]*yt.ex5.3.r2 + e1[201:400]
yt.ex5.2.r2 <- -beta.ex5.r2.2[2]*yt.ex5.3.r2 + e2[201:400]
yt.ex5.r2 <- cbind(yt.ex5.1.r2, yt.ex5.2.r2, yt.ex5.3.r2)

## model construction: regime 3 
yt.ex5.2.r3 <- sum(e2[1:400]) + cumsum(e2[401:600])
yt.ex5.3.r3 <- sum(e3[1:400]) + cumsum(e3[401:600])
yt.ex5.1.r3 <- -beta.ex5.r3[2]*yt.ex5.2.r3 -beta.ex5.r3[3]*yt.ex5.3.r3 + e1[401:600]
yt.ex5.r3 <- cbind(yt.ex5.1.r3, yt.ex5.2.r3, yt.ex5.3.r3)

yt.ex5 <- rbind(yt.ex5.r1, yt.ex5.r2, yt.ex5.r3)

ts.plot(yt.ex5, lty = c(1:d), 
        main = "Trivariate With Cointegration Regime Changes", 
        col = rainbow(d))
legend(0, -20, c("y1","y2", "y3"), lty=c(1:3), col = rainbow(d))
```

```{r, EXAMPLE 5 COINTEGRATION CHECK PART 1}
## test for presence of unit roots; all processes nonstationary 
adf.test(yt.ex5[,1], k = 1)
adf.test(yt.ex5[,2], k = 1)
adf.test(yt.ex5[,3], k = 1)

## verify all I(1) processes; note result is iffy due to regime changes...
adf.test(diff(yt.ex5[,1]), k = 1)
adf.test(diff(yt.ex5[,2]), k = 1)
adf.test(diff(yt.ex5[,3]), k = 1)

## confirm VAR order  
p.ex5 <- VARselect(yt.ex5)
p.ex5$selection

## trace and eigenvalue statistic using urca -> ca.jo() 
ex5.eigen.stat <- ca.jo(yt.ex5, type = "eigen", ecdet = "none", K = 2, 
                        spec = "transitory")
summary(ex5.eigen.stat)

ex5.trace.stat <- ca.jo(yt.ex5, type = "trace", ecdet = "none", K = 2, 
                        spec = "transitory")
summary(ex5.trace.stat)
```

With extremely sharp regime switching, trace and eigenvalue tests fail to detect cointegration. 

```{r, EXAMPLE 5 COINTEGRATION CHECK PART 2}
## rolling window helper function  
window.min.max <- function(num.index, win.length, roll.length) {
  init.val <- (num.index - win.length) / roll.length
  window.hold <- floor(win.length / roll.length) - 1
  min.window <- floor(init.val) + 1
  max.window <- min.window + window.hold
  list(min.window, max.window)
}

win.len = 100; roll.len = 5

## data cleaning for windowing
yt.ex5.window <- yt.ex5 %>% 
  as.data.frame() %>%
  mutate(num.index = 0:(n-1)) %>% 
  mutate(window.min = window.min.max(num.index, win.len, roll.len)[[1]], 
         window.max = window.min.max(num.index, win.len, roll.len)[[2]])

roll.window <- function(data) {
  max.window <- max(data$window.max)
  cr.vec <- rep(NA, max.window); counter <- 0
  
  for (i in 0:(max.window-2)) {
    temp.dat <- data %>% filter(window.min <= i & window.max >= i)
    temp.eig <- ca.jo(temp.dat[,1:3], K = 2,type = "eigen", ecdet = "none", 
                      spec = "transitory")
    if (counter == 0) {cval <- temp.eig@cval[,2]; counter <- 1}
    test.stat <- temp.eig@teststat
    coint.rank <- sum(test.stat > cval)
    cr.vec[i+1] <- coint.rank
  }
  cr.vec
}

coint.rank <- roll.window(yt.ex5.window)
index.min <- seq(1, 595, by = 5)
index.max <- seq(101, 695, by = 5)
cr.ind <- cbind(coint.rank, index.min, index.max)
data.frame(cr.ind)
```

Cointegration rank seems extremely sensitive -- overlapping with the shifting regime in any capacity derails cointegration and leads to cointegration rank zero (albeit regime shift is extremely sharp, which likely breaks relation down very quickly). Accurate cointegration rank testing only possible if we know what regime we're actually in. 

```{r, EXAMPLE 5 EIGEN ANGLES, fig.align="center", out.width="50%"}
## calculate changing angle between cointegration vector across roll windows 
angle <- function(v1, v2) {
  theta <- acos(sum(v1*v2) / (norm(v1, type = "2")*norm(v2, type = "2")))
  theta
}

## angle calculation holding cointegration rank static at 1 
roll.angle.1 <- function(data, c.rank = 1) {
  max.window <- max(data$window.max) 
  angle.vec <- rep(NA, max.window)
  
  init.dat <- data %>% filter(window.min <= 0 & window.max >= 0)
  init.vecm <- VECM(init.dat[,1:3], r = c.rank, lag = 1, estim = "ML", include = "const")
  prev.beta <- coefB(init.vecm)
  
  for (i in 1:(max.window-2)) {
    temp.dat <- data %>% filter(window.min <= i & window.max >= i)
    temp.vecm <- VECM(temp.dat[,1:3], r = c.rank, 
                          lag = 1, estim = "ML", include = "const")
    temp.beta <- coefB(temp.vecm)
    temp.angle <- angle(temp.beta, prev.beta)
    angle.vec[i] <- temp.angle
    prev.beta <- temp.beta
  }
  angle.vec 
}

## angle calculation with cointegration rank 2; use projection for selection 
proj.min.2d <- function(v1, v2, w1, w2) {
  pairs <- list(cbind(v1,w1), cbind(v1,w2), cbind(v2,w1), cbind(v2,w2))
  curr.min <- 1e6
  pair.min <- NA
  for (pair in pairs) {
    temp.len <- sum(pair[,1]*pair[,2]) / norm(pair[,2], type = "2")
    if (temp.len < curr.min) {
      pair.min <- pair
    }
  }
  pair.min 
}

roll.angle.2 <- function(data, c.rank = 2) {
  max.window <- max(data$window.max) 
  angle.vec <- rep(NA, max.window)
  
  init.dat <- data %>% filter(window.min <= 0 & window.max >= 0)
  init.vecm <- VECM(init.dat[,1:3], r = c.rank, lag = 1, estim = "ML", include = "const")
  prev.beta <- coefB(init.vecm)
  
  for (i in 1:(max.window-2)) {
    temp.dat <- data %>% filter(window.min <= i & window.max >= i)
    temp.vecm <- VECM(temp.dat[,1:3], r = c.rank, 
                          lag = 1, estim = "ML", include = "const")
    temp.beta <- coefB(temp.vecm)
    proj.pair <- proj.min.2d(temp.beta[,1], temp.beta[,2], 
                             prev.beta[,1], prev.beta[,2])
    temp.angle <- angle(proj.pair[,1], proj.pair[,2])
    angle.vec[i] <- temp.angle
    prev.beta <- temp.beta
  }
  angle.vec 
}

angle.r1 <- roll.angle.1(yt.ex5.window, 1)
angle.r2 <- roll.angle.2(yt.ex5.window, 2)

par(mfrow=c(1,2))
ts.plot(na.omit(angle.r1), main = "Angle Between CI Basis: Rank 1", 
        xlab = "Window", ylab = "Angle (degrees)")
ts.plot(na.omit(angle.r2), main = "Angle Between CI Basis: Rank 2", 
        xlab = "Window", ylab = "Angle (degrees)")
```

Clear separation in angle measurement regions when holding rank 1 cointegration constant; difficult to establish clear separation of regimes via angle measures when holding rank 2 cointegration constant. 

```{r, EXAMPLE 6 / EXTENDED EX 5 EIGEN ANGLES KNOWN REGIMES}
## check angle measures in cointegration without regime switching: example 3
n <- length(yt.ex3[,1])

yt.ex3.window <- yt.ex3 %>% 
  as.data.frame() %>%
  mutate(num.index = 0:(n-1)) %>% 
  mutate(window.min = window.min.max(num.index, win.len, roll.len)[[1]], 
         window.max = window.min.max(num.index, win.len, roll.len)[[2]])

ex3.angle.1 <- roll.angle.1(yt.ex3.window)
ex3.angle.2 <- roll.angle.2(yt.ex3.window)

par(mfrow=c(1,2))
ts.plot(ex3.angle.1, main = "EX3 Angles, CI Rank 1", 
        xlab = "Window", ylab = "Angle (degrees)")
ts.plot(ex3.angle.2, main = "EX3 Angles, CI Rank 2", 
        xlab = "Window", ylab = "Angle (degrees)")
## check angle measures in cointegration without regime switching: example 4
n <- length(yt.ex4[,1])

yt.ex4.window <- yt.ex4 %>% 
  as.data.frame() %>%
  mutate(num.index = 0:(n-1)) %>% 
  mutate(window.min = window.min.max(num.index, win.len, roll.len)[[1]], 
         window.max = window.min.max(num.index, win.len, roll.len)[[2]])

ex4.angle.1 <- roll.angle.1(yt.ex4.window)
ex4.angle.2 <- roll.angle.2(yt.ex4.window)
par(mfrow=c(1,2))
ts.plot(ex4.angle.1, main = "EX4 Angles, CI Rank 1", 
        xlab = "Window", ylab = "Angle (degrees)")
ts.plot(ex4.angle.2, main = "EX4 Angles, CI Rank 2", 
        xlab = "Window", ylab = "Angle (degrees)")
```

```{r, EXAMPLE 6 VERIFICATION}
## verify roll.window() accurately assesses cointegration rank for EX3, EX4
(ex3.windowed.cr <- roll.window(yt.ex3.window))
ex3.prop <- sum(na.omit(ex3.windowed.cr) == 1) / length(na.omit(ex3.windowed.cr))
  
(ex4.windowed.cr <- roll.window(yt.ex4.window))
ex4.prop <- sum(na.omit(ex4.windowed.cr) == 2) / length(na.omit(ex4.windowed.cr))

ex3.prop; ex4.prop
```

Example 3 and example 4 cointegration rank correctly determined by 88.78% and 94.90% of tests. 